\Large{All in the family:\\German twin finishing times in\\the 2016 women's Olympic marathon}

### 2017-02-04 paper/paper.tex
edits

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio women's marathon varied as a function
of athletes' personal best times and as function of their half
marathon split times.  The points in the plots are colored by
twin/triplet status, and both plots contain second-order polynomial
regression lines.  Rug marks in both plot indicate times (either
personal best or half marathon split) of runners who earned a DNF
result.

@@

Considering first the relationship between Olympic finishing and
personal best times, Figure \ref{fig:45degreeplot}'s solid 45-degree
line is informative.  Given the paucity of points (only five of them)
below this line, it follows that the vast majority of Olympic
marathoners ran slower in Rio compared to prior personal bests.
Relative to personal best times, the Hahner twins were definitely on
the slow side, but a number of runners had greater differences between
their Olympic times and their personal bests than the Hahners.  These
runners are denoted with squares in Figure \ref{fig:45degreeplot}, and
there are 13 such symbols, highlighting approximately 10\% of the
racers who completed the Rio marathon.  Moreover, the figure shows
that there were two women who had personal best times slightly faster
than the Hahner's and yet finished after the German women.  Although
Figure \ref{fig:45degreeplot} suggests that the Hahner twins were
slower than one would have expected given their previous best marathon
times, it is not consistent with the accusation that they dramatically
slowed down in the Rio marathon.

@@

With respect to our second measure of marathon skill, Figure
\ref{fig:45degreeplot_half} shows that there was nothing abnormal
about the Hahner twins' overall finishing times, conditional on their
half marathon split times.  This is consistent with the previous
Figure \ref{fig:45degreeplot} and inconsistent with accusations made
against the Hahner twins, at least the part of the accusation that
focused on their overall pace.

Figure \ref{fig:residuals} is a visualization of how unusual, or not,
the Hahner and Kim twins' finishing times were in the Olympic
marathon, conditional on our two measures of marathon talent.  The
left and right panels of the figure are based on the polynomial
regressions in Figure \ref{fig:scatter}'s panels, respectively.  We
calculate residuals from Figures \ref{fig:45degreeplot} and
\ref{fig:45degreeplot_half}, we Studentize the residuals, and then we
plot the ordered residuals in Figure \ref{fig:residuals}.

@@

The cumulative residuals depicted in Figure \ref{fig:residuals} are
consistent with our interpretation of the Hahner twin's Olympic
finishing times as not particularly remarkable.  With personal best
time as a measure of marathon talent, the two Hahner residuals are in
the right tail of the residual distribution; however, their locations
are not extreme.  One residual is located at the 88th percentile and
the other at the 93rd.  Kim twin residuals are similarly not
remarkable although they are in the left tail of the residual
distribution, indicating that the Kims ran faster than one would have
expected.  If one uses half marathon split times as a measure of
marathon talent, similar results obtain.  Neither Hahner twin had a
finishing time that was particularly unusual given her half marathon
split, and this applies to the Kim twins as well.

@@

The personal best times of the Hahner twins were 115 seconds apart and
their official finishing times were separated by one second.  Is such
a 115 to one compression typical among pairs of runners?  Moreover,
are there other pairs of Olympic marathoners who had a difference
between personal best times of 115 seconds apart and, if so, how close
where their finishing times?

@@

Thinking about the accusations leveled against the Hahner twins,
Figure \ref{fig:diffdiffscatter} suggests that Anna and Lisa Hahner
did indeed run with an eye on each other.  In fact, the same can be
said of the Kim twins, who ran seemingly in lockstep throughout the
entire Rio marathon.  The North Korean twins had a personal best
difference of 53 seconds and a finishing time difference of literally
zero seconds.  Beyond these twins, there were eight pairs of Rio
runners with a 53 second personal best difference, and resulting
finishing time differences are as follows: 9, 51, 228, 340, 352, 571,
662, and 751.  As in the Hahner case, the Kim twins compressed their
finishing times more than any other pair of runners with a similar
personal best difference.

>>>>

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio women's marathon varied as a function
of athletes' personal best times and as function of their half
marathon split times.  The points in the plots are colored by
twin/triplet status, and both plots contain second-order polynomial
regression lines.  The ticks along the x-axis in both plots indicate times (either
personal best or half marathon split) of runners who earned a DNF
result.

@@

Considering first the relationship between Olympic finishing and
personal best times, Figure \ref{fig:45degreeplot}'s solid 45-degree
line is informative.  Given the paucity of points (only five of them)
below this line, it follows that the vast majority of Olympic
marathoners ran slower in Rio compared to their prior personal bests.
The Hahner twins were definitely on the slow side, well above the 45-degree line,
but a number of runners had even greater differences between
their Olympic times and their personal bests than the Hahners.  These
runners are denoted with squares in Figure \ref{fig:45degreeplot}, and
there are 13 such symbols, highlighting approximately 10\% of the
racers who completed the Rio marathon.  Moreover, the figure shows
that there were two women who had personal best times slightly faster
than the Hahner's and yet finished after the German women.  Although
Figure \ref{fig:45degreeplot} suggests that the Hahner twins were
slower than one would have expected given their previous best marathon
times, it is not consistent with the accusation that they dramatically
slowed down in the Rio marathon.

@@

With respect to our second measure of marathon skill, Figure
\ref{fig:45degreeplot_half} shows that there was nothing abnormal
about the Hahner twins' overall finishing times, conditional on their
half marathon split times. As one might expect, a runner's time halfway through the race is a
fairly strong predictor of their finishing time at the end of the race.  Here we see that the relationship
between the twins' halfway time and their finishing time is similar to that
of the other runners.  The two points representing the Hahners are
in the middle of the distribution of points and therefore imply that the twins do not seem to have slowed down in any unusual way. 
 This is consistent with the previous
Figure \ref{fig:45degreeplot} and inconsistent with accusations made
against the Hahner twins, at least the part of the accusation that
focused on their overall pace.

Figure \ref{fig:residuals} gives a visualization of how unusual, or not,
the Hahner and Kim twins' finishing times were in the Olympic
marathon, conditional on our two measures of marathon talent.  The
left and right panels of the figure display the studentized residuals derived from the polynomial
regressions in Figure \ref{fig:scatter}'s panels, respectively.  Hence, we
compute the residuals from Figures \ref{fig:45degreeplot} and
\ref{fig:45degreeplot_half}, Studentize them, and arrange the result from least to most
along the x-axis in Figure \ref{fig:residuals} to display their cumulative distribution.

@@

The cumulative residuals depicted in Figure \ref{fig:residuals} are
consistent with our interpretation of the Hahner twin's Olympic
finishing times as not particularly remarkable.  With personal best
time as a measure of marathon talent, the two Hahner residuals are in
the right tail of the residual distribution; however, their locations
are not extreme.  One residual is located at the 88th percentile and
the other at the 93rd.  While they are in the tail end of the distribution, they are not the
major outliers that would lead us to reject the twins' claim.  
Moreover, the residuals for the Kim twins are similarly not
remarkable, although they are in the left tail of the residual
distribution, indicating that the Kims ran faster than one would have
expected.  If one uses half marathon split times as a measure of
marathon talent, similar results obtain.  Neither Hahner twin had a
finishing time that was particularly unusual given her half marathon
split, and this applies to the Kim twins as well.

@@

The personal best times of the Hahner twins were 115 seconds apart and
their official finishing times were separated by one second.  Is such
a 115 to one compression typical among pairs of runners?  Moreover,
are there other pairs of Olympic marathoners who had a difference
between personal best times of 115 seconds apart and, if so, how close
were their finishing times?

@@

Thinking about the accusations leveled against the Hahner twins,
Figure \ref{fig:diffdiffscatter} suggests that Anna and Lisa Hahner
did indeed run with an eye on each other.  In fact, the same can be
said of the Kim twins, who ran seemingly in lockstep throughout the
entire Rio marathon.  The North Korean twins had a personal best
difference of 53 seconds and a finishing time difference of literally
zero seconds.  Beyond these twins, there were eight pairs of Rio
runners with a 53 second personal best difference, and resulting
finishing time differences are as follows: 9, 51, 228, 340, 352, 571,
662, and 751.  As in the Hahner case, the Kim twins compressed their
finishing times - meaning that they finished with less time between  
them than the difference in their personal bests - more than any other pair of runners with a similar
personal best difference.
